### 2023-09-04 proofs/tex/norm-lemmas.tex
WIP: leftovers
WIP: description, notation
Grammar update: Γ = set

\begin{lemma}
  \label{lemma:fv-nf}
  Free variables are not changed by the normalization
  \begin{itemize}
  \item[$-$] $[[fv iN]] \equiv [[fv nf(iN)]]$
  \item[$+$] $[[fv iP]] \equiv [[fv nf(iP)]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  By straightforward induction on $[[nf(iN) = iM]]$.
\end{proof}

@@

      $\begin{aligned}[t]
         [[ nf([σ]iN) ]] &= [[ nf([σ] ∀pas.iM)]] \\
                          &= [[ nf(∀pas.[σ]iM) ]]
                          && \text{Assuming $[[{pas} ∩ {Γ1}]] = \emptyset$
                             and $[[{pas} ∩ {Γ2}]] = \emptyset$}\\
                          &= [[ ∀pbs.nf([σ]iM) ]]
                          && \text {Where $[[pbs]] = [[ord {pas} in nf([σ]iM)]]
                             = [[ord {pas} in [σ]iM]]$ (the latter is by
                             \cref{corollary:normalization-ord})}\\
                          &= [[ ∀pas'.nf([σ]iM) ]]
                          && \text{By \cref{lemma:ord-sigma}, $[[pbs]] = [[pas']]$
                             since $[[{pas}]]$ is disjoint with $[[Γ1]]$ and
                             $[[Γ2]]$}\\
                          &= [[ ∀pas'.[nf(σ)]nf(iM) ]]
                          && \text {By the induction hypothesis}\\
         \end{aligned}$ \\

     To show alpha-equivalence of 
     $[[ [nf(σ)] ∀pas'.nf(iM) ]]$ and $[[ ∀pas'.[nf(σ)]nf(iM) ]]$,
     we can assume that $[[{pas'} ∩ {Γ1}]] = \emptyset$, and $[[{pas'} ∩ {Γ2}]]
     = \emptyset$.

@@

  Suppose that $[[σ]]$ is a substitution  $[[Γ2 ⊢ σ : Γ1]]$,
  $[[iP]]$ is a positive type ($[[Γ1 ⊢ iP]]$),
  $[[iN]]$ is a negative type ($[[Γ1 ⊢ iN]]$). Then
  \begin{itemize}
  \item[$+$]
    $[[ [σ]iP  ]] \text{ is normal} \iff
    \begin{cases}
      [[σ|fv(iP)]] &\text{is normal}\\
      [[iP]]       &\text{is normal}\\
    \end{cases} $
  \item[$-$]
    $[[ [σ]iN  ]] \text{is normal} \iff
    \begin{cases}
      [[σ|fv(iN)]] &\text{is normal}\\
      [[iN]]       &\text{is normal}\\
    \end{cases} $
  \end{itemize}
\end{lemma}

\begin{proof}
  Mutual induction on $[[Γ1 ⊢ iP]]$ and $[[Γ1 ⊢ iN]]$.
  \begin{caseof}
  \item $[[iN]] = [[na]]$\\
    Then $[[iN]]$ is always normal, and
    the normality of $[[σ|{na}]]$ by the definition means $[[ [σ]na ]]$ is normal.

@@

                                           && \text{by congruence of normality \ilyam{lemma?}}\\

@@

                                          \iff&&%

>>>>

\begin{lemma}
  \label{lemma:fv-nf}
  Free variables are not changed by the normalization
  \begin{itemize}
  \item[$-$] $[[fv iN]] \equiv [[fv nf(iN)]]$
  \item[$+$] $[[fv iP]] \equiv [[fv nf(iP)]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  By straightforward induction on $[[iN]]$ and mutually on $[[iP]]$.
\end{proof}

\begin{lemma}
  \label{lemma:uv-nf}
  Algorithmic variables are not changed by the normalization
  \begin{itemize}
  \item[$-$] $[[uv uN]] \equiv [[uv nf(uN)]]$
  \item[$+$] $[[uv uP]] \equiv [[uv nf(uP)]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  By straightforward induction on $[[uN]]$ and mutually on $[[uP]]$.
\end{proof}

@@

      $\begin{aligned}[t]
         [[ nf([σ]iN) ]] &= [[ nf([σ] ∀pas.iM)]] \\
                          &= [[ nf(∀pas.[σ]iM) ]]
                          && \text{Assuming $[[{pas} ∩ Γ1]] = \emptyset$
                             and $[[{pas} ∩ Γ2]] = \emptyset$}\\
                          &= [[ ∀pbs.nf([σ]iM) ]]
                          && \text {Where $[[pbs]] = [[ord {pas} in nf([σ]iM)]]
                             = [[ord {pas} in [σ]iM]]$ (the latter is by
                             \cref{corollary:normalization-ord})}\\
                          &= [[ ∀pas'.nf([σ]iM) ]]
                          && \text{By \cref{lemma:ord-sigma}, $[[pbs]] = [[pas']]$
                             since $[[{pas}]]$ is disjoint with $[[Γ1]]$ and
                             $[[Γ2]]$}\\
                          &= [[ ∀pas'.[nf(σ)]nf(iM) ]]
                          && \text {By the induction hypothesis}\\
         \end{aligned}$ \\

     To show alpha-equivalence of 
     $[[ [nf(σ)] ∀pas'.nf(iM) ]]$ and $[[ ∀pas'.[nf(σ)]nf(iM) ]]$,
     we can assume that $[[{pas'} ∩ Γ1]] = \emptyset$, and $[[{pas'} ∩ Γ2]]
     = \emptyset$.

@@

  Suppose that $[[σ]]$ is a substitution  $[[Γ2 ⊢ σ : Γ1]]$,
  $[[iP]]$ is a positive type ($[[Γ1 ⊢ iP]]$),
  $[[iN]]$ is a negative type ($[[Γ1 ⊢ iN]]$). Then
  \begin{itemize}
  \item[$+$]
    $[[ [σ]iP  ]] \text{ is normal} \iff
    \begin{cases}
      [[σ|fv(iP)]] &\text{is normal}\\
      [[iP]]       &\text{is normal}\\
    \end{cases} $
  \item[$-$]
    $[[ [σ]iN  ]] \text{is normal} \iff
    \begin{cases}
      [[σ|fv(iN)]] &\text{is normal}\\
      [[iN]]       &\text{is normal}\\
    \end{cases} $
  \end{itemize}
\end{lemma}
\begin{proof}
  Mutual induction on $[[Γ1 ⊢ iP]]$ and $[[Γ1 ⊢ iN]]$.
  \begin{caseof}
  \item $[[iN]] = [[na]]$\\
    Then $[[iN]]$ is always normal, and
    the normality of $[[σ|{na}]]$ by the definition means $[[ [σ]na ]]$ is normal.

@@

                                             \end{cases}\\

@@

                                             \end{cases}\\

### 2023-09-04 proofs/tex/ord-lemmas.tex
WIP: leftovers
Grammar update: Γ = set

@@

\begin{lemma}[Ordering is not affected by independent substitutions]
  \label{lemma:ord-sigma}
  Suppose that $[[Γ2 ⊢ σ : Γ1]]$, i.e. $[[σ]]$ maps variables from $[[Γ1]]$ into types
  taking free variables from $[[Γ2]]$, and $[[varset]]$ is a set of variables
  disjoint with both $[[Γ1]]$ and $[[Γ2]]$, 
  $[[iN]]$ and $[[iP]]$ are types. Then
    \begin{itemize}
  \item[$-$] $[[ ord varset in [σ]iN ]] = [[ord varset in iN ]]$
  \item[$+$] $[[ ord varset in [σ]iP ]] = [[ord varset in iP ]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  Mutual induction on $[[iN]]$ and $[[iP]]$.
  \begin{caseof}
    \item $[[iN = na]]$ \\
      If $[[na ∉ {Γ1}]]$ then $[[ [σ]na = na ]]$ and $[[ ord varset in [σ]na ]] = [[ ord varset in na ]]$, 
      as requried.
      If $[[na ∊ {Γ1}]]$ then $[[na ∉ varset]]$, so $[[ ord varset in na ]] = [[·]]$.
      Moreover, $[[Γ2 ⊢ σ : Γ1]]$ means $[[ fv([σ]na) ⊆ {Γ2} ]]$, and thus, 
      as a set, $[[ ord varset in [σ]na ]] = [[varset ∩ fv([σ]na)]] \subseteq [[varset ∩ {Γ2}]] = [[·]]$.
    \item $[[iN = ∀pas.iM]]$\\
      We can assume $[[{pas} ∩ {Γ1} = ∅]]$
      and $[[{pas} ∩ varset = ∅]]$. Then

>>>>

\begin{corollary}[Idempotency of ordering]
  \label{corollary:ord-idemp}
  \hfill
  \begin{itemize}
    \item[$-$] If $[[ ord varset in iN = ordVars ]]$ then 
      $[[ ord {ordVars} in iN = ordVars ]]$,
    \item[$+$] If $[[ ord varset in iP = ordVars ]]$ then 
      $[[ ord {ordVars} in iP = ordVars ]]$;
  \end{itemize}
\end{corollary}
\begin{proof}
  By \cref{lemma:ord-soundness,corollary:ord-weakening}.
\end{proof}

\end{proof}

@@

\begin{lemma}[Ordering is not affected by independent substitutions]
  \label{lemma:ord-sigma}
  Suppose that $[[Γ2 ⊢ σ : Γ1]]$, i.e. $[[σ]]$ maps variables from $[[Γ1]]$ into types
  taking free variables from $[[Γ2]]$, and $[[varset]]$ is a set of variables
  disjoint with both $[[Γ1]]$ and $[[Γ2]]$, 
  $[[iN]]$ and $[[iP]]$ are types. Then
    \begin{itemize}
  \item[$-$] $[[ ord varset in [σ]iN ]] = [[ord varset in iN ]]$
  \item[$+$] $[[ ord varset in [σ]iP ]] = [[ord varset in iP ]]$
  \end{itemize}
\end{lemma}
\begin{proof}
  Mutual induction on $[[iN]]$ and $[[iP]]$.
  \begin{caseof}
    \item $[[iN = na]]$ \\
      If $[[na ∉ Γ1]]$ then $[[ [σ]na = na ]]$ and $[[ ord varset in [σ]na ]] = [[ ord varset in na ]]$, 
      as requried.
      If $[[na ∊ Γ1]]$ then $[[na ∉ varset]]$, so $[[ ord varset in na ]] = [[·]]$.
      Moreover, $[[Γ2 ⊢ σ : Γ1]]$ means $[[ fv([σ]na) ⊆ Γ2 ]]$, and thus, 
      as a set, $[[ ord varset in [σ]na ]] = [[varset ∩ fv([σ]na)]] \subseteq [[varset ∩ Γ2]] = [[·]]$.
    \item $[[iN = ∀pas.iM]]$\\
      We can assume $[[{pas} ∩ Γ1 = ∅]]$
      and $[[{pas} ∩ varset = ∅]]$. Then
